Ensembler: Enabling high-throughput molecular simulations at the superfamily scale

### 2015-03-18 manuscript/ms.tex
Moved residue numbering schemes into an appendix in the main text

@@

To evaluate the models of \emph{Src} and \emph{Abl1} in the context of the published literature, we have focused on two residue pair distances thought to be important for the regulation of protein kinase domains.
We use the residue numbering schemes for chicken Src (which is commonly used in the literature even in reference to human Src)[CITE: 2SRC, 1Y57] and human Abl1 isoform A[CITE: 2F4J, 2HYY, 2G1T] respectively; the exact numbering schemes are provided in Supporting Information S1.
Fig.~\ref{figure:src-ref-structures} shows two structures of \emph{Src} believed to represent inactive (PDB code: 2SRC) {\color{red}[CITE: 2SRC]} and active (PDB code: 1Y57) {\color{red}[CITE: 1Y57]} states.
One notable feature which distinguishes the two structures is the transfer of an electrostatic interaction of E310 from R409 (in the inactive state) to K295 (in the active state), brought about by a rotation of the $\alpha$C-helix.
These three residues are also well conserved [CITE Kannan Neuwald JMB 2005] and a number of experimental and simulation studies have suggested that this electrostatic switching process plays a role in a regulatory mechanism shared across the protein kinase family [CITE Foda Shan Seeliger Src Nat Commun 2015; Shukla Pande Nat Commun 2014; Ozkirimli Post Prot Sci 2008].
As such, we have projected the {\bf Ensembler} models for \emph{Src} and \emph{Abl1} onto a space consisting of the distances between these two residue pairs (Fig.~\ref{figure:pair-distances}).
The models show strong coverage of regions in which either of the electrostatic interactions is formed, as well as a wide range of regions inbetween.
We thus expect that such a set of models, if used as starting configurations for highly parallel MD simulation, could greatly aid in sampling of the activation process.

@@

>>>>

\usepackage{alltt}

@@

To evaluate the models of \emph{Src} and \emph{Abl1} in the context of the published literature, we have focused on two residue pair distances thought to be important for the regulation of protein kinase domains.
We use the residue numbering schemes for chicken Src (which is commonly used in the literature even in reference to human Src)[CITE: 2SRC, 1Y57] and human Abl1 isoform A[CITE: 2F4J, 2HYY, 2G1T] respectively; the exact numbering schemes are provided in Supporting Information S\ref{si:residue-numbering-schemes}.
Fig.~\ref{figure:src-ref-structures} shows two structures of \emph{Src} believed to represent inactive (PDB code: 2SRC) {\color{red}[CITE: 2SRC]} and active (PDB code: 1Y57) {\color{red}[CITE: 1Y57]} states.
One notable feature which distinguishes the two structures is the transfer of an electrostatic interaction of E310 from R409 (in the inactive state) to K295 (in the active state), brought about by a rotation of the $\alpha$C-helix.
These three residues are also well conserved [CITE Kannan Neuwald JMB 2005] and a number of experimental and simulation studies have suggested that this electrostatic switching process plays a role in a regulatory mechanism shared across the protein kinase family [CITE Foda Shan Seeliger Src Nat Commun 2015; Shukla Pande Nat Commun 2014; Ozkirimli Post Prot Sci 2008].
As such, we have projected the {\bf Ensembler} models for \emph{Src} and \emph{Abl1} onto a space consisting of the distances between these two residue pairs (Fig.~\ref{figure:pair-distances}).
The models show strong coverage of regions in which either of the electrostatic interactions is formed, as well as a wide range of regions inbetween.
We thus expect that such a set of models, if used as starting configurations for highly parallel MD simulation, could greatly aid in sampling of the activation process.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% APPENDICES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\onecolumngrid
\newpage
\appendix

\renewcommand{\thesection}{\arabic{section}}
\section{Sequences and residue numbering schemes for Src and Abl1}
\label{si:residue-numbering-schemes}

Kinase catalytic domains are highlighted in red, and the conserved residues analyzed in the main text (Figs.~\ref{figure:src-ref-structures} and~\ref{figure:pair-distances}) are highlighted with yellow background.

\subsubsection*{Human Abl1 sequence}

\begin{alltt}
1     MLEICLKLVG CKSKKGLSSS SSCYLEEALQ RPVASDFEPQ GLSEAARWNS KENLLAGPSE    60
61    NDPNLFVALY DFVASGDNTL SITKGEKLRV LGYNHNGEWC EAQTKNGQGW VPSNYITPVN   120
121   SLEKHSWYHG PVSRNAAEYL LSSGINGSFL VRESESSPGQ RSISLRYEGR VYHYRINTAS   180
181   DGKLYVSSES RFNTLAELVH HHSTVADGLI TTLHYPAPKR NKPTVYGVSP NYDKWEMERT   240
241   D{\color{red}ITMKHKLGG GQYGEVYEGV WKKYSLTVAV {\setlength{\fboxsep}{0pt}\colorbox{yellow}{K}}TLKEDTMEV EEFLK{\setlength{\fboxsep}{0pt}\colorbox{yellow}{E}}AAVM KEIKHPNLVQ}   300
301   {\color{red}LLGVCTREPP FYIITEFMTY GNLLDYLREC NRQEVNAVVL LYMATQISSA MEYLEKKNFI}   360
361   {\color{red}HRDLAARNCL VGENHLVKVA DFGLS{\setlength{\fboxsep}{0pt}\colorbox{yellow}{R}}LMTG DTYTAHAGAK FPIKWTAPES LAYNKFSIKS}   420
421   {\color{red}DVWAFGVLLW EIATYGMSPY PGIDLSQVYE LLEKDYRMER PEGCPEKVYE LMRACWQWNP}   480
481   {\color{red}SDRPSFAEIH QAF}ETMFQES SISDEVEKEL GKQGVRGAVS TLLQAPELPT KTRTSRRAAE   540
541   HRDTTDVPEM PHSKGQGESD PLDHEPAVSP LLPRKERGPP EGGLNEDERL LPKDKKTNLF   600
601   SALIKKKKKT APTPPKRSSS FREMDGQPER RGAGEEEGRD ISNGALAFTP LDTADPAKSP   660
661   KPSNGAGVPN GALRESGGSG FRSPHLWKKS STLTSSRLAT GEEEGGGSSS KRFLRSCSAS   720
721   CVPHGAKDTE WRSVTLPRDL QSTGRQFDSS TFGGHKSEKP ALPRKRAGEN RSDQVTRGTV   780
781   TPPPRLVKKN EEAADEVFKD IMESSPGSSP PNLTPKPLRR QVTVAPASGL PHKEEAGKGS   840
841   ALGTPAAAEP VTPTSKAGSG APGGTSKGPA EESRVRRHKH SSESPGRDKG KLSRLKPAPP   900
901   PPPAASAGKA GGKPSQSPSQ EAAGEAVLGA KTKATSLVDA VNSDAAKPSQ PGEGLKKPVL   960
961   PATPKPQSAK PSGTPISPAP VPSTLPSASS ALAGDQPSST AFIPLISTRV SLRKTRQPPE  1020
1021  RIASGAITKG VVLDSTEALC LAISRNSEQM ASHSAVLEAG KNLYTFCVSY VDSIQQMRNK  1080
1081  FAFREAINKL ENNLRELQIC PATAGSGPAA TQDFSKLLSS VKEISDIVQR             1130
\end{alltt}

\subsubsection*{Sequences for human and chicken Src, aligned using Clustal Omega}

\begin{alltt}
SRC_HUMAN   1     MGSNKSKPKD ASQRRRSLEP AENVHGAGGG AFPASQTPSK PASADGHRGP SAAFAPAAAE    60
SRC_CHICK   1     MGSSKSKPKD PSQRRRSLEP PDSTH---HG GFPASQTPNK TAAPDTHRTP SRSFGTVATE    57
                  ***.******  *********  :..*    * .*******.*  *: * ** * * :*. .*:*  
SRC_HUMAN   61    PKLFGGFNSS DTVTSPQRAG PLAGGVTTFV ALYDYESRTE TDLSFKKGER LQIVNNTEGD   120
SRC_CHICK   58    PKLFGGFNTS DTVTSPQRAG ALAGGVTTFV ALYDYESRTE TDLSFKKGER LQIVNNTEGD   117
                  ********:* **********  ********* ********** ********** **********  
SRC_HUMAN   121   WWLAHSLSTG QTGYIPSNYV APSDSIQAEE WYFGKITRRE SERLLLNAEN PRGTFLVRES   180
SRC_CHICK   118   WWLAHSLTTG QTGYIPSNYV APSDSIQAEE WYFGKITRRE SERLLLNPEN PRGTFLVRES   177
                  *******:** ********** ********** ********** ******* ** **********  
SRC_HUMAN   181   ETTKGAYCLS VSDFDNAKGL NVKHYKIRKL DSGGFYITSR TQFNSLQQLV AYYSKHADGL   240
SRC_CHICK   178   ETTKGAYCLS VSDFDNAKGL NVKHYKIRKL DSGGFYITSR TQFSSLQQLV AYYSKHADGL   237
                  ********** ********** ********** ********** ***.****** **********  
SRC_HUMAN   241   CHRLTTVCPT SKPQTQGLAK DAWEIPRES{\color{red}L RLEVKLGQGC FGEVWMGTWN GTTRVAI{\setlength{\fboxsep}{0pt}\colorbox{yellow}{K}}TL}   300
SRC_CHICK   238   CHRLTNVCPT SKPQTQGLAK DAWEIPRES{\color{red}L RLEVKLGQGC FGEVWMGTWN GTTRVAI{\setlength{\fboxsep}{0pt}\colorbox{yellow}{K}}TL}   297
                  *****.**** ********** ********** ********** ********** **********  
SRC_HUMAN   301   {\color{red}KPGTMSPEAF LQ{\setlength{\fboxsep}{0pt}\colorbox{yellow}{E}}AQVMKKL RHEKLVQLYA VVSEEPIYIV TEYMSKGSLL DFLKGETGKY}   360
SRC_CHICK   298   {\color{red}KPGTMSPEAF LQ{\setlength{\fboxsep}{0pt}\colorbox{yellow}{E}}AQVMKKL RHEKLVQLYA VVSEEPIYIV TEYMSKGSLL DFLKGEMGKY}   357
                  ********** ********** ********** ********** ********** ****** ***  
SRC_HUMAN   361   {\color{red}LRLPQLVDMA AQIASGMAYV ERMNYVHRDL RAANILVGEN LVCKVADFGL A{\setlength{\fboxsep}{0pt}\colorbox{yellow}{R}}LIEDNEYT}   420
SRC_CHICK   358   {\color{red}LRLPQLVDMA AQIASGMAYV ERMNYVHRDL RAANILVGEN LVCKVADFGL A{\setlength{\fboxsep}{0pt}\colorbox{yellow}{R}}LIEDNEYT}   417
                  ********** ********** ********** ********** ********** **********  
SRC_HUMAN   421   {\color{red}ARQGAKFPIK WTAPEAALYG RFTIKSDVWS FGILLTELTT KGRVPYPGMV NREVLDQVER}   480
SRC_CHICK   418   {\color{red}ARQGAKFPIK WTAPEAALYG RFTIKSDVWS FGILLTELTT KGRVPYPGMV NREVLDQVER}   477
                  ********** ********** ********** ********** ********** **********  
SRC_HUMAN   481   {\color{red}GYRMPCPPEC PESLHDLMCQ CWRKEPEERP TFEYLQAFLE DYF}TSTEPQY QPGENL       536
SRC_CHICK   478   {\color{red}GYRMPCPPEC PESLHDLMCQ CWRKDPEERP TFEYLQAFLE DYF}TSTEPQY QPGENL       533
                  ********** ********** ****:***** ********** ********** ******
\end{alltt}
